Classless Java

### 2015-11-21 paper/sections/Overview.tex
More on Section 2

%\lstinputlisting[linerange=46-52]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=POINT2D

@@

%Now we proceed to define \texttt{Animal} with \texttt{point} ``member
%field''.
%%Not again, before we used only getters!
% Again, we model this member field with getter and setter methods:
\lstinputlisting[linerange=64-67]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=ANIMAL

@@

\lstinputlisting[linerange=71-74]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=HORSE

@@

@@

\subsubsection*{\Q@Point3D@ and properties updater}
\Q@Bird@ needs a 3D location, thus first of all we define \Q@Point3D@ and
as follows:
\lstinputlisting[linerange=56-60]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=POINT3D
\marco{Fix over: with(Point must be with(Point2D
and it should be after withZ, and on more lines}
Just to give you a feeling on how much boring code \mixin is generating,  we show the expanded code without then \mixin annotation\footnote{
As you can see, the generated code is very repetitive, writing such code by hand can easily induce bugs, for example a distract programmer may
swap the arguments of one of the many calls of \Q@Point3D.of@.
}.\begin{lstlisting}
interface Point3D extends Point2D{
  Point3D withX(int val);    Point3D withY(int val);    Point3D withZ(int val);
  Point3D with(Point2D val);
  public static Point3D of(int _x, int _y, int _z){
    int x=_x;    int y=_y;    int z=_z;
    return new Point3D(){
      public int x(){return x;}    public int y(){return y;}    public int z(){return z;}
      public Point3D withX(int val){return Point3D.of(val,this.y(),this.z());}
      public Point3D withY(int val){return Point3D.of(this.x(),val,this.z());}
      public Point3D withZ(int val){return Point3D.of(this.x(),this.y(),val);}
      public Point3D with(Point2D val){
        if(val instanceof Point3D){return (Point3D)val;}
        return Point3D.of(val.x(),val.y(),this.z());
        }      };    }  }
\end{lstlisting}
Note how \Q@with-@ methods are automatically refined in their return type, so that code like 
\Q@Point3D p=Point3D.of(1,2,3); p=p.withX(42);@ will be accepted and behave as expected. If the programmer wishes to suppress this behaviour and keep the signature as it was, it is sufficient to redefine the \Q@with-@ methods in the new class repeating the old signature.
Again, the philosophy is that if the programmer provides something directly, \mixin do not touch it.
\Q@with-@ methods (functionally) update a field/property at a time.
This can be inefficient, and sometime hard to maintain.
Often we want to update many fields at the same time, for example using another object as source.
Following this idea, the method \Q@with(Point2D)@ is an example of a (functional) properties updater: 
it takes a certain type and in the current object update all field in that type that match fields in the current type.
The idea is that we want as result something that is still like \Q@this@, but modified to be as much as possible similar to the parameter.
The cast in \Q@with(Point2D)@ is trivially safe since it is guarded by an \Q@instanceof@ test. The idea is that if the parameter is a subtype of the current exact type, then we can just return the parameter, as something that is just ``more'' than \Q@this@.

@@

\lstinputlisting[linerange=78-86]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=BIRD

@@

\subsection{\Q@Pegasus@ and multiple Inheritance}
The \emph{``multiple inheritance''} case appears at interface
\texttt{Pegasus}.
Pegasus (one of the best known creatures
in Greek mythology) can not only \emph{run} but also \emph{fly}! 
It would be handy if they could obtain \texttt{fly} and \texttt{run} functionality from both \texttt{Horse} and \texttt{Bird}.

@@

\lstinputlisting[linerange=94-95]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=PEGASUS

@@

\lstinputlisting[linerange=46-52]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=POINT2D

\lstinputlisting[linerange=56-60]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=POINT3D

\lstinputlisting[linerange=64-67]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=ANIMAL

\lstinputlisting[linerange=71-74]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=HORSE

\lstinputlisting[linerange=78-86]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=BIRD

\lstinputlisting[linerange=94-95]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=PEGASUS

>>>>

%\lstinputlisting[linerange=50-56]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=POINT2D

@@

%Now we proceed to define \texttt{Animal} with \texttt{point} ``member
%field''.
%%Not again, before we used only getters!
% Again, we model this member field with getter and setter methods:
\lstinputlisting[linerange=68-71]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=ANIMAL

@@

\lstinputlisting[linerange=75-78]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=HORSE

\noindent Creating and using a horse animal is quite simple:

\lstinputlisting[linerange=9-11]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=USINGHORSE

\noindent Note how the \texttt{of}, \texttt{withX} and
\texttt{location} methods (all generated automatically) provide a
basic interface for dealing with animals.

\paragraph{Summary} Dealing with state (whether mutable or immutable)
in object interfaces relies on a notion of abstract state, where only
methods to interact with state are available to users. Object
interfaces provides support for four different types abstract state
operations, which are summarized in Figure~\ref{fig:abstractstate}.
The abstract state operations are determined by naming conventions and the
types of the methods. Fluent setters are a variant of mutable setters,
and are discussed in the case study in Section~\ref{}.

\begin{figure}
\begin{tabular}{|l|l|l|}
\hline
 & \textbf{Example}                  & \textbf{Description } \\ \hline
\textbf{``fields''/getters}        &   \Q@int x();@                  & Retrieves value from field \texttt{x}.          \\ \hline
{\bf withers}        &   \Q@Point2D withX(int val);@                & Clones
object; updates field \texttt{x} to \texttt{val}.             \\ \hline
\textbf{mutable setters}        & \Q@void x(int val);@ & Sets the field
\texttt{x} to a  new value \texttt{val}.        \\ \hline
\textbf{fluent setters}        & \Q@Point2D x(int val);@ &Sets the field
\texttt{x} to \texttt{val} and returns \texttt{\this}.           \\ \hline
\end{tabular}

\caption{Abstract state operations, for a field \texttt{x}, allowed by the \mixin
  annotation. }

\label{fig:abstractstate}

\end{figure}

@@

\subsection{Object Interfaces and Subtyping}

@@

\subsubsection*{\Q@Point3D@ and properties updater}
\Q@Bird@ needs a 3D location, thus first of all we define \Q@Point3D@ and
as follows:
\lstinputlisting[linerange=60-64]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=POINT3D
\marco{Fix over: with(Point must be with(Point2D
and it should be after withZ, and on more lines}
Just to give you a feeling on how much boring code \mixin is generating,  we show the expanded code without then \mixin annotation\footnote{
As you can see, the generated code is very repetitive, writing such code by hand can easily induce bugs, for example a distract programmer may
swap the arguments of one of the many calls of \Q@Point3D.of@.
}.\begin{lstlisting}
interface Point3D extends Point2D{
  Point3D withX(int val);    Point3D withY(int val);    Point3D withZ(int val);
  Point3D with(Point2D val);
  public static Point3D of(int _x, int _y, int _z){
    int x=_x;    int y=_y;    int z=_z;
    return new Point3D(){
      public int x(){return x;}    public int y(){return y;}    public int z(){return z;}
      public Point3D withX(int val){return Point3D.of(val,this.y(),this.z());}
      public Point3D withY(int val){return Point3D.of(this.x(),val,this.z());}
      public Point3D withZ(int val){return Point3D.of(this.x(),this.y(),val);}
      public Point3D with(Point2D val){
        if(val instanceof Point3D){return (Point3D)val;}
        return Point3D.of(val.x(),val.y(),this.z());
        }      };    }  }
\end{lstlisting}
Note how \Q@with-@ methods are automatically refined in their return type, so that code like 
\Q@Point3D p=Point3D.of(1,2,3); p=p.withX(42);@ will be accepted and behave as expected. If the programmer wishes to suppress this behaviour and keep the signature as it was, it is sufficient to redefine the \Q@with-@ methods in the new class repeating the old signature.
Again, the philosophy is that if the programmer provides something directly, \mixin do not touch it.
\Q@with-@ methods (functionally) update a field/property at a time.
This can be inefficient, and sometime hard to maintain.
Often we want to update many fields at the same time, for example using another object as source.
Following this idea, the method \Q@with(Point2D)@ is an example of a (functional) properties updater: 
it takes a certain type and in the current object update all field in that type that match fields in the current type.
The idea is that we want as result something that is still like \Q@this@, but modified to be as much as possible similar to the parameter.
The cast in \Q@with(Point2D)@ is trivially safe since it is guarded by an \Q@instanceof@ test. The idea is that if the parameter is a subtype of the current exact type, then we can just return the parameter, as something that is just ``more'' than \Q@this@.

@@

\lstinputlisting[linerange=82-90]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=BIRD

@@

\subsection{Advanced Multiple Inheritance with Object Interfaces}
The \emph{``multiple inheritance''} case appears at interface
\texttt{Pegasus}.
Pegasus (one of the best known creatures
in Greek mythology) can not only \emph{run} but also \emph{fly}! 
It would be handy if they could obtain \texttt{fly} and \texttt{run} functionality from both \texttt{Horse} and \texttt{Bird}.

@@

\lstinputlisting[linerange=98-99]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=PEGASUS

@@

\lstinputlisting[linerange=50-56]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=POINT2D

\lstinputlisting[linerange=60-64]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=POINT3D

\lstinputlisting[linerange=68-71]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=ANIMAL

\lstinputlisting[linerange=75-78]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=HORSE

\lstinputlisting[linerange=82-90]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=BIRD

\lstinputlisting[linerange=98-99]{../UseMixinLombok/src/test/TestAnimal.java}% APPLY:linerange=PEGASUS
